### 2017-12-11 CO681-QuantumInformationProcessing/project/network-flows.tex
minima finding

\subsection{Directed Graph Layering}
\paragraph{}
An important subproblem we will use in solving Network Flow problems, and again later in solving Matching problems is the computing of a {\it layering}. Given a connected directed graph $G = (V,E)$ and a source vertex $r \in V$ we are to find a labelling, which we call a layering, on the vertices, $\ell : V \rightarrow \N$ such that $\ell(r) = 0$ and for all $v\neq r$
$$ \ell(v) = 1+ \min_{u:(u,v) \in E} \ell(u).$$
\paragraph{Algorithm}
We can compute a layering as follows:
\begin{enumerate}
\item Initialize $\ell(r) = 0$ and $\ell(v) = \infty$ for all $v \neq r$. Further initialize a queue $W = \{r\}$.
\item While $W\neq \emptyset$
	\begin{enumerate}
	\item Pop $v$ off the head of the queue $W$
	\item Find by Grover's Search all neighbours $w$ with $\ell(w) = \infty$.
	\item Set $\ell(w) = \ell(v) + 1$ and append $w$ to $W$.
	\end{enumerate} 
\end{enumerate}
\begin{theorem}\label{th:layering}
The algorithm above computes a layering for $G=(V,E)$  in time $O(n^{3/2}\log n)$ in the adjacency model and in time $O(\sqrt{nm}\log n)$ in the list model, where $n = |V|$ and $m = |E|$.
\end{theorem}
\begin{proof}
Clearly initialization takes time $O(n)$, and every vertex is processed at most once. In the adjacency model the processing of every vertex takes $O(\sqrt{n})$ by Theorem \ref{th:grovers}. The log-factor appears in accordance with Note \ref{note:log}.
\paragraph{}
In the list model, each vertex is processed in time $O(\sqrt{\frac{n_v}{d(v)}}d_v + \sqrt{d(v) + 1}) = O(\sqrt{n_vd(v)} + \sqrt{d_v + 1})$, where $d(v)$ is the out-degree of $v$, and $n_v$ is the number of neighbours of $v$ with $\ell(v)=\infty$ when $v$ is searched from. Using the Cauchy-Schwarz inequality we can bound the expected running time for all vertices by
$$\sum_{v\in V} \sqrt{n_vd(v)} \leq \sqrt{\sum_{v\in V} n_v}\sqrt{\sum_{v \in V} d(v)} = O(\sqrt{nm}),$$
and similarly for $\sum_{v \in V} \sqrt{d(v) + 1}$. Hence the claimed running time.
\end{proof}
\subsection{Integer Network Flow Algorithm}

>>>>

\subsection{Directed Graph Layering}
\paragraph{}
An important subproblem we will use in solving Network Flow problems, and again later in solving Matching problems is the computing of a {\it layering}. Given a connected directed graph $G = (V,E)$ and a source vertex $r \in V$ we are to find a labelling, which we call a layering, on the vertices, $\ell : V \rightarrow \N$ such that $\ell(r) = 0$ and for all $v\neq r$
$$ \ell(v) = 1+ \min_{u:(u,v) \in E} \ell(u).$$
\paragraph{Algorithm}
We can compute a layering as follows:
\begin{enumerate}
\item Initialize $\ell(r) = 0$ and $\ell(v) = \infty$ for all $v \neq r$. Further initialize a queue $W = \{r\}$.
\item While $W\neq \emptyset$
	\begin{enumerate}
	\item Pop $v$ off the head of the queue $W$
	\item Find by Grover's Search all neighbours $w$ with $\ell(w) = \infty$.
	\item Set $\ell(w) = \ell(v) + 1$ and append $w$ to $W$.
	\end{enumerate} 
\end{enumerate}
\begin{theorem}\label{th:layering}
The algorithm above computes a layering for $G=(V,E)$  in time $O(n^{3/2}\log n)$ in the adjacency model and in time $O(\sqrt{nm}\log n)$ in the list model, where $n = |V|$ and $m = |E|$.
\end{theorem}
\begin{proof}
Clearly initialization takes time $O(n)$, and every vertex is processed at most once. In the adjacency model the processing of every vertex takes $O(\sqrt{n})$ by Theorem \ref{th:grovers}. The log-factor appears in accordance with Note \ref{note:log}.
\paragraph{}
In the list model, each vertex is processed in time $$O(\sqrt{\frac{n_v}{d(v)}}d_v + \sqrt{d(v) + 1}) = O(\sqrt{n_vd(v)} + \sqrt{d_v + 1})$$ where $d(v)$ is the out-degree of $v$, and $n_v$ is the number of neighbours of $v$ with $\ell(v)=\infty$ when $v$ is searched from. Using the Cauchy-Schwarz inequality we can bound the expected running time for all vertices by
$$\sum_{v\in V} \sqrt{n_vd(v)} \leq \sqrt{\sum_{v\in V} n_v}\sqrt{\sum_{v \in V} d(v)} = O(\sqrt{nm}),$$
and similarly for $\sum_{v \in V} \sqrt{d(v) + 1}$. Hence the claimed running time.
\end{proof}
\subsection{Integer Network Flow Algorithm}

### 2017-12-11 CO681-QuantumInformationProcessing/project/quantum-primitives.tex
minima finding

>>>>

We present a notion of the sort of indices we want to add to $I$, split into two subtypes. For any $j \in \{1, \dots, N\}$, we call $j$ {\it good for} $I$ if there exists some $i \in I$ such that $f(j) < f(i)$ and either
\item {\it(Known Type)} $g(j) = g(i)$, or 
\item {\it(Unknown Type)} $g(j) \not\in g(I)$.
We present the following algorithm for minima finding:
\item Initialize $I$ as a set of $d$ dummy indices of unique type with respect to $g$ and unique maximal value with respect to $f$
\item Repeat until no good elements for $I$
	\item Use Grover's Search to find a good element $j$ for $I$.
	\item If $j$ is of Known type, replace $i \in I$ such that $f(j) < f(i)$ and $g(j) = g(i)$ with $j$
	\item Otherwise (i.e. if $j$ is of Unknown type) let $i \in I$ be element of largest $f$-value. Replace $i$ with $j$.
\begin{lemma}\label{lemma:minima}Let $I \subseteq \{1,\dots, N\}$ such that $|I| = d$ and for all $i,j \in I$, $g(i) \neq g(j)$ with $t>0$ good elements for $I$. Then after $O(d)$ iterations of step $2$, in expectation, there are at most $\frac{3}{4}t$ good elements for $I$.
\end{lemma}
See Lemma $4$ of D{\"u}rr et al. \cite{durr2004quantum}.
\begin{theorem}\label{th:minima}
The algorithm above solves the minima finding problem in expected time $O(\sqrt{dN})$.
Let $t$ denote the number of good elements remaining for $I$ at a given iteration of the algorithm. From Lemma \ref{lemma:minima} and Theorem \ref{th:grovers}, we have that, if $t > 2d$ then there are at most $\frac{t}{2}$ good elements for $I$ after $O(d\sqrt{\frac{N}{t}})$ queries of $f$ in expectation. Hence for $t>2d$ the expected number of queries of $f$ until $t\leq 2d$ is
$$O\big(d\big(\sqrt{\frac{N}{d}} + \sqrt{\frac{N}{2d}} + \sqrt{\frac{N}{4d}} + \dots\big)\big) = O(\sqrt{dN}).$$
Then once the algorithm has $t \leq 2d$, the number of queries of $f$ needed in expectation to reach $t=0$ is
$$O\big(\sum_{j=1}^{2d}\sqrt{\frac{N}{j}}\big) = O(\sqrt{dN})$$
as desired.
Two special case of minima finding are the problems of Finding $d$ Smallest Values of a Function, and Find $d$ Elements of Different Type. By a reduction from distinguishing matrices to these special subproblems, they show in \cite{durr2004quantum} that Minima Finding requires $\Omega(\sqrt{dN})$ queries of $f$ and hence the algorithm we present is optimal.
